Below-ground sink limitation alters growth and carbon balance of Eucalyptus seedlings

### 2015-10-26 manuscript/manuscript_pve.Rmd
made a new table of root data
edits from MT comments

RLD and SRL need to be double checked

leafN is no without TNC and is in manuscript

#table data
library(pixiedust)
table1 <- read.csv("master_scripts/pve_table1.csv")
table2 <- read.csv("master_scripts/pve_table2.csv")
modeltable <- read.csv("master_scripts/modeltable.csv")

@@

## Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed abd the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a 25 g subsample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Here, specific root length (SRL) is reported as the root length divided by the dry mass of each subsample (m g^-1^).

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^), for leaves sampled during gas exchange campaigns, was then calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass.

@@

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but only in some of the container size treatments (Table 1). Over the duration of the experiment SLA~f~ was higher in free seedlings, but was not different across containers sizes (Table 1, P < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (P < 0.001).

## Leaf chemistry
Leaf N % was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001).  Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N % compared to other soil volumes, while free seedlings maintained the highest leaf N % (Table 1, P < 0.001).  Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table 1).  Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001).

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 2) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at muliple bin levels (n=5) of the covariate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was postively correlated with final harvest biomass across all seedlings (P < 0.001).

\
Both J~max~ and Vc~max~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table 2). Leaf dark respiration rates were not significantly different across soil volumes (Table 2). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table 2).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although g~s~ in free seedlings was generally higher than those in containers (Table 2. P < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table 1). Combined these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N % at harvest was not different across soil volumes ($\bar{x}$ = 4.5 %), with minimal decreases from pre-planting value ($\bar{x}$ = 4.9 %). This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

\
Here, reductions in A were postively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. Physical root restriction or decreased supply as been shown cause reductions in N uptake (citations), which could explain the observed reductions in leaf N in the smallest containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead faciliated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N % declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the affects of belowground sink limitation on A of seedlings in containers was evidents throughout the experiment.

@@

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may prevent further growth [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Using measured LMF from the harvest, instead of the optimized seedling control (M~2~), improved biomass predictions and provided insight into how sink limitation can impact leaf C allocation beyond A. The sensitivity of the model to shifts in LMF could represent leaf loss throughout the experiment that could not be explicitly quantified in this field study. As TNC accumulation can lead to accelerated leaf senescence [@paul2001sink], this could explain the observed decline in total leaf area of seedlings in small containers. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of A and the fate of C allocated to growth, respiration and C storage in leaves. It will be the interactions between these two components that will determine the total C gain available for plant growth.

\
Increasing rates of non-leaf respiration (M~3~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. This is noteworthy as it shows that sink limitation could differentially affect respiration of component tissues. Additionally, keeping rates of respiration fixed in growth models may underestimate impacts of environmental change on tissue respiration above and belowground. The fraction of photosynthate used in respiration varies substantially among species and environments  and is sensitive to changes in growth rates [@Lambers2008]. As C balance is a delicate equilibrium between fluxes of A and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions (Valentini et al. 2000). We agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond A, is likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate a growth will undoubtedly shift across different experimental manipulations and plant species.

@@

**Table 1**. Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the free seedling treatment.

\
**Table 2.** Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. A~max~, R~dark~ and g~s~ are each measured at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n=6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

# Tables {.unnumbered}
**Table 1**.Responses of plant and leaf characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

dust(table1)+
  sprinkle_colnames("Volume (L)", 
                        "Seedling mass (g)", 
                        "SLA~TNC-free~ (m^2^ kg^-1^)",
                        "Leaf Starch (%)", 
                        "Leaf Sugars (%)", 
                        "Leaf Nitrogen (%)",
                        "Root Nitrogen (%)",
                        "SRL (cm m^-1^)",
                        paste0("Leaf ", greeks['delta'],"^13^C"," (\u2030)")) +
  sprinkle(rows= 1, part="head",bold=TRUE)+
  sprinkle_print_method("markdown")
```

\
**Table 2.** Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. A~max~, R~dark~ and g~s~ are each measured at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n=6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

```{r table2, echo=FALSE, warning=FALSE}
dust(table2)+
  sprinkle_colnames("Volume (L)", 
                        "A~max~ ($\\mu$mol m^-2^ s^-1^)", 
                        "R~dark~ ($\\mu$mol m^-2^ s^-1^)",
                        "J~max~",
                        "Vc~max~",
                        "g~s~ (mol m^-2^ s^-1^)",
                        "g~1~") +
  sprinkle(rows= 1, part="head",bold=TRUE)+
  sprinkle_print_method("markdown")
```

@@

**Figure S1**. Seedling growth model default parameters.
```{r tabels1, echo=FALSE, warning=FALSE}
dust(modeltable)+
  sprinkle(rows= 1, part="head",bold=TRUE)+
  sprinkle_print_method("markdown")
```

>>>>

#table data
library(pixiedust)
table1 <- read.csv("master_scripts/pve_table1.csv")
table2 <- read.csv("master_scripts/root_table.csv")
table3 <- read.csv("master_scripts/pve_table2.csv")
modeltable <- read.csv("master_scripts/modeltable.csv")

@@

## Growth and morphology metrics
Seedlings were planted in summer (January 21^st^ 2013) and stem height, diameter at 15 cm and leaf count were measured weekly thereafter. Once the growth rate of individual plants had significantly declined a full biomass harvest was completed abd the experiment ended (May 21^st^ 2013). Dry mass of leaves, stems, roots and total leaf area (LI-3100C Area Meter; LI-COR, Lincoln, NE, USA) were measured for each seedling. Mean individual leaf area for each harvested seedling was calculated by dividing total measured leaf area by total leaf count of only fully expanded leaves. Mean individual leaf area was then used to interpolate total seedling leaf area through time with weekly leaf counts. Root mass was collected by removing the roots system and passing soil from each container through a 1 mm sieve, washing, separating into fine and coarse roots (<2 mm and >2 mm diameter, respectively) and then drying to a constant mass. Roots of seedlings in containers were not considered pot bound, as clusters of roots along the soil-container interface were not observed. Roots from the free seedlings were collected by excavating each 1 m^2^ subplot to the hard layer and keeping only roots within the subplot. For each seedling, a subsample of washed fine roots was analyzed for root length using WhinoRhizo software (Regent Instruments Inc., Quebec, QC, Canada). Specific root length (SRL) is reported as the root length divided by the dry mass of each subsample (m g^-1^). Root length density (RLD) is reported as the root length divided by the total volume of each container (m dm^-3^).

@@

\
Leaf total non-structural carbohydrate (TNC) concentration was analyzed on dried and milled leaf samples using a total starch assay kit (Megazyme International, Wicklow, Ireland) and includes the starch  and soluble sugar concentrations (mg g^-1^). Starch was quantified using a thermostable $\alpha$-amylase and amyloglucosidase assay [@McCleary_starch] and soluble sugars were determined following the anthrone method [@ebell1969variation]. Complete methods of the TNC assay are described in [@mitchell2013drought]. TNC-free specific leaf area (SLA~f~, m^2^ kg^-1^) for leaves sampled during gas exchange campaigns, were calculated by first subtracting the TNC content from individual dry leaf mass before dividing leaf area by leaf mass or the leaf nitrogen content. Similarily, TNC-free leaf N (N~f~, %) was calculated on all gas exchange leaves from leaf mass content without TNC and leaf N content.

@@

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but only in some of the container size treatments (Table 2). Root length density increased  (Table 2)

Over the duration of the experiment SLA~f~ was higher in free seedlings, but was not different across containers sizes (Table 1, P < 0.001) and this pattern was evident beginning in the first gas exchange measurement campaign (P < 0.001).

## Leaf and root chemistry
Leaf N~f~ was significantly higher in free seedlings and the largest container volume compared to the smaller container volumes at the onset of gas exchange measurements (6th week, P < 0.001). Throughout the remainder of the experiment the smallest container volume had a significant reduction in leaf N~f~ compared to other soil volumes, while free seedlings maintained the highest leaf N~f~ (Table 1, P < 0.001). Leaf starch content in the smallest container was ca. double  that of free seedlings (P = 0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table 1). Differences in leaf starch between the free seedling and the smallest container were evident during the first gas exchange campaign (P = 0.001). Root N % was higher in free seedlings compared to seedlings in containers but only for some of the container size treatments (Table 2).

## Gas exchange and photosynthetic parameters
At the first measurement campaign, both A~sat~ and A~max~ were significantly higher in the free seedling treatment compared to seedlings in containers (both P < 0.001). Across all measurement campaigns mean A~sat~ (Figure `r asatfig`) and A~max~ (Table 3) were consistently higher in free seedlings than in containers (26 % and 29 %, respectively). The relationships between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ on a mass basis was highly correlated to both leaf N content and leaf starch (both P < 0.001). We used predictions from the linear mixed effect model equation to visualize these relationship of A~max~ to either leaf N content or leaf starch at muliple bin levels (n=5) of the covariate parameter (Figure `r achemfig`). Across all measurement campaigns and treatments A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b). Overall, A~max~ was postively correlated with final harvest biomass across all seedlings (P < 0.001).

\
Both J~max~ and Vc~max~ were significantly higher in free seedlings (30 % and 26 %, respectively) than container-grown seedlings with little variation between container volume treatments (Table 3). Leaf dark respiration rates were not significantly different across soil volumes (Table 3). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2011reconciling] optimal stomatal conductance model, was lowest in the free seedling treatment and was marginally different across soil volume treatments (Table 3).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 MPa across all seedlings, respectively. Although g~s~ in free seedlings was generally higher than those in containers (Table 3. P < 0.001), the mean rates for all seedlings were high at 0.37 mol m^-2^ s^-1^ and did not change throughout the course of the experiment. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table 1). Combined these indices provide strong evidence that water stress was not apparent on these seedlings throughout the experiment. Soil N % at harvest was not different across soil volumes ($\bar{x}$ = 4.5 %), with minimal decreases from pre-planting value ($\bar{x}$ = 4.9 %). This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

@@

\
Here, reductions in A were postively correlated with decreases in leaf N and leaf N was considerably reduced for seedlings in containers. As leaf N reductions were detected with TNC-free leaf mass, this suggests that physical root restriction or decreased supply likely affected seedling N uptake instead of TNC dilution in leaves in the smalles containers. Root N at the end of the experiment was on average higher in free seedlings but not consistently higher than every container volume treatment. Unrestricted mycorrhizal recruitment could have instead faciliated the increases in leaf N in free seedlings, but this effect is unknown. Combined with the fact that soil N % declined evenly across all treatments, there was no clear mechanism present to identify changes in root N uptake between free seedlings and seedlings in containers. In these already low quality soils, it is possible that seedlings in containers simply grew into increasing N limitation which negatively affected belowground sink strength. Although no clear feedback could be determined between the available soil N pool and decreases in leaf N, the affects of belowground sink limitation on A of seedlings in containers was evidents throughout the experiment.

@@

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. The sensitivity of roots to their own exudates near obstructions may prevent further growth [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments. This is not surprising as plants in containers have been shown to have different root morphology to field grown plants [@nesmith1998effect]. The poor soil quality used in our experiment and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Using measured LMF from the harvest, instead of the optimized seedling control (M~s1~), improved biomass predictions and provided insight into how sink limitation can impact leaf C allocation beyond A. The sensitivity of the model to shifts in LMF could represent leaf loss throughout the experiment that could not be explicitly quantified in this field study. As TNC accumulation can lead to accelerated leaf senescence [@paul2001sink], this could explain the observed decline in total leaf area of seedlings in small containers. Future empirical and modelling studies should focus on how feedbacks from sink activity affect both rates of A and the fate of C allocated to growth, respiration and C storage in leaves. It will be the interactions between these two components that will determine the total C gain available for plant growth.

\
Increasing rates of non-leaf respiration (M~s2~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. We also show that very large increases in non-leaf respiration rates would have been required to accurately predict observed seedling biomass. Although the fraction of photosynthate used in respiration is known to vary among species and is sensitive to changes in growth rates [@Lambers2008], results from M~s2~ highlight a lack of knowledge regarding how respiration rates of individual tissues, within a single plant, maybe be differentially affected by environmental change. This is noteworthy, as C balance is a delicate equilibrium between fluxes of A and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions (Valentini et al. 2000). These results from infer that using fixed rates of respiration in models likely underestimates plant responses to sink limitation. Thus, we agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond A, are likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate a growth will undoubtedly shift across different experimental manipulations and plant species.

@@

**Table 1**.Responses of plant and leaf characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

\
**Table 2.** Responses of root characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. All values are from the final harvest. Vaules for RLD are only calculated for seedlings in containers as free seedlings have potentially unlimeted soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for RLD which represents only diffeences between seedlings in containers.

\
**Table 3.** Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. A~max~, R~dark~ and g~s~ are each measured at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n=6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

# Tables {.unnumbered}
**Table 1**.Responses of plant and leaf characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass and leaf $\delta$^13^C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

@@

dust(table1)%>%
  sprinkle_colnames("Volume (L)", 
                        "Seedling mass (g)", 
                        "SLA~TNC-free~ (m^2^ kg^-1^)",
                        "Leaf Starch (%)", 
                        "Leaf Sugars (%)", 
                        "Leaf Nitrogen~TNC-free~ (%)",
                        paste0("Leaf ", greeks['delta'],"^13^C"," (\u2030)")) %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

**Table 2.** Responses of root characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. All values are from the final harvest. Vaules for RLD were only calculated for seedlings in containers as free seedlings had potentially unlimited soil volume to exploit. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings, except for RLD which represents only diffeences between seedlings in containers.
```{r, echo=FALSE, warning=FALSE}
dust(table2)%>%
  sprinkle_colnames("Volume (L)", 
                        "Root Nitrogen (%)", 
                        "SRL (m g^-1^)",
                        "RLD (m dm^-3^)") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

\
**Table 3.** Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. A~max~, R~dark~ and g~s~ are each measured at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n=6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The volume effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

```{r table2, echo=FALSE, warning=FALSE}
dust(table3)%>%
  sprinkle_colnames("Volume (L)", 
                        "A~max~ ($\\mu$mol m^-2^ s^-1^)", 
                        "R~dark~ ($\\mu$mol m^-2^ s^-1^)",
                        "J~max~",
                        "Vc~max~",
                        "g~s~ (mol m^-2^ s^-1^)",
                        "g~1~") %>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```

@@

**Figure S1**. Seedling growth model default parameters.
```{r tabels1, echo=FALSE, warning=FALSE}
dust(modeltable)%>%
  sprinkle(rows= 1, part="head",bold=TRUE)%>%
  sprinkle_print_method("markdown")
```
